Global patterns of forest autotrophic carbon fluxes

### 2020-05-11 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd

\newpage
###Abstract
Carbon fixation, allocation, and metabolism by trees set the basis for energy and material flows in forest ecosystems and define their interactions with Earth's changing climate. 
Yet, we lack a cohesive synthesis on how forest autotrophic carbon fluxes (FACF) vary globally with respect to climate and one another. 
Here, we draw upon # records from the Global Forest Carbon Database (ForC), representing all major forest types and the nine most significant FACF, to comprehensively explore how C cycling in mature, undisturbed forests varies with latitude and climate on a global scale. 
We show that, across all FACF variables analyzed, C cycling decreases linearly with latitude -- a finding that confirms multiple previous studies but contradicts the idea that net primary productivity ($NPP$) of temperate forests rivals that of tropical forests. 
FACF generally displayed similar trends acrooss latitude and multiple climate variables, with few differences in allocation detectable at this global scale, but differed in that latitude explained a lower proportion of variation among subsidiary fluxes (in particular, woody aboveground $NPP$ and belowground $NPP$, $BNPP$). 
Climate explained a significant proportion (#-#%) of variation in all C fluxes analyzed, with temperature variables in general and mean annual temperature (MAT) in particular being the best predictors of FACF on this global scale. 
The effects of temperature were modified by moisture availability, with FACF reduced under hot and dry conditions and sometimes under very high precipitaiton. 
FACF declined with temperature seasonality, but growing season length did not improve upon MAT as a predictor.  
Within the growing season, the influence of climate on C cycling was small but significant for a number of FACF. 
These findings clarify the big picture of how FACF varies with latitude and climate on a global scale. 
As we enter a period of accelerating climatic change, understanding of the fundamental climatic controls on FACF sets a foundation for understanding patterns of change.

@@

Carbon cycling in forests worldwide provides the energetic basis for sustaining the majority of Earth's terrestrial biodiversity and many human populations (**REF**), while strongly influencing atmospheric CO~2~ and climate [@bonan_forests_2008]. Forests' autotrophic carbon fluxes (FACF)--that is, carbon fixation, allocation, and metabolism by trees and other primary producers--sets the energy ultimately available to heterotrophic organisms (including microbes), in turn influencing their abundance (e.g., [Zak et al. 1994](http://esajournals.onlinelibrary.wiley.com/doi/abs/10.2307/1940888); [Niedziałkowska et al. 2010](https://doi.org/10.2980/17-1-3310)) and possibly diversity [@waide_relationship_1999; Chu et al. 2019 (DOI: 10.1111/ele.13175)]. FACF influences all organic matter stocks in forest ecosystems and is linked to cycling of energy, water, and nutrients. Critically, FACF also define forest interactions with Earth's changing climate. Over 69 Gt of CO~2~ cycle through Earth's forests each year [@badgley_terrestrial_2019], and in recent decades their net C sequestration (~2.4 Gt C yr^-1^) offset roughly 30% of anthropogenic fossil fuel emissions [@pan_large_2011]. As atmospheric carbon dioxide levels continue to rise, driving climate change, forests will play a critical role in shaping the future of Earth's climate [@cavaleri_urgent_2015; @rogelj_mitigation_2018]. However, our ability to draw general macroscopic conclusions regarding global variation in FACF with respect to climate has been limited in that these analyses often mix forests that vary in stand age, disturbance history, and/or management status; do not always sufficiently parse related variables; and typically consider only one or a few variables at a time.

FACF decrease with latitude, but it remains unclear whether and how the shape of this relationship varies across fluxes. Studies agree that FACF are lowest in the boreal regions, and increase into the temperate regions [@luyssaert_co_2007; @huston_global_2009; @beer_terrestrial_2010; @piao_forest_2010; @jung_global_2011]. However, evidence is inconclusive on whether primary productivity continues to increase into the tropics, or whether it plateaus in temperate regions. Evidence for this is further complicated by the fact that different studies use different measures of productivity to explore these relationships. For example, modelling of global terrestrial ecosystem gross primary productivity ($GPP$) through upscaling and calibration of eddy flux measurements indicates that GPP peaks in tropical forests [@beer_terrestrial_2010; @jung_global_2011; @badgley_terrestrial_2019; @li_mapping_2019]. In contrast, some studies suggest that the highest values of net primary productivity ($NPP$) may be found in temperate forests [@luyssaert_co_2007; @huston_global_2009], while others find $NPP$ highest in the tropics and decreasing with latitude [@simova_enigma_2017]. Other studies have chosen to focus exclusively on above-ground net primary productivity ($ANPP$), finding evidence of a weak negative relationship between $ANPP$ and latitude [@huston_global_2009; @gillman_latitude_2015].

The latitudinal gradient in FACF is primarily driven by climate, which is a significant driver of FACF across broad spatial scales [@luyssaert_co_2007; @cleveland_relationships_2011; @hursh_sensitivity_2017]. The majority of studies have focused on exploring the relationships of FACF to mean annual temperature (MAT) and mean annual precipitation (MAP), as the most commonly reported site-level climate variables. While these fail to capture some important aspects of climate such as seasonality, they do describe broad trends in temperature and water availability, and therefore capture a substantial portion of global-scale variation in climate. There is strong evidence that both MAT and MAP show significant positive relationships with FACF [@chu_does_2016]. However, as with latitude, the shape of those relationships is not always clear, and, again, is complicated by the use of different measures of FACF across studies. Various measures of primary productivity {**FACF?**} saturate at high levels of MAP, though the saturation points identified vary from 1500mm [@luyssaert_co_2007] up to 2445mm MAP [@schuur_productivity_2003]. Studies of the influence of MAT on productivity {**FACF?**} are less conclusive. Luyssaert et al. [-@luyssaert_co_2007] examined GPP and NPP and found that, while $GPP$ increases linearly with MAT, NPP saturates at around 10$^\circ$C MAT. In contrast, Larjavaara and Muller-Landau [-@larjavaara_temperature_2012], find that increases in $GPP$ saturate at approximately 25$^\circ$C MAT, while Schuur [-@schuur_productivity_2003] shows that $NPP$ increases linearly with temperature. Taylor et al. [-@taylor_temperature_2017] showed a positive interaction between MAT and MAP in shaping tropical forest productivity, such that high rainfall had a negative effect on productivity in cooler climates, compared to a positive effect in warmer climates. {**It would be good to add some more citations on soil respiraiton. I'm sure BBL can help.**}

FACF can be influenced by many other factors as well, which often act across a range of scales, and may show interactive effects with each other [@cleveland_relationships_2011]. On a local scale, stand age [@litton_carbon_2007; @gillman_latitude_2015], biodiversity [@liang_positive_2016], management [@simova_enigma_2017], nutrient availability [@aragao_above_2009], and altitude [@girardin_net_2010; @malhi_variation_2017] all impact FACF. On a global scale, we expect that FACF are most strongly influenced by broad climatic gradients. There is evidence that FACFs also respond to variables such as cloud cover [@taylor_temperature_2017], solar radiation [@fyllas_solar_2017], and potential evapotranspiration [@kerkhoff_plant_2005] in potentially significant ways. Furthermore, MAT and MAP are very coarse measures of climate, and so fail to capture much variation in climate on an intra-annual scale, including the effects of factors such as growing season length, number of frost-free days, temperature seasonality, and dry season length. Some studies have suggested that the apparently strong relationship between MAT and FACFs is actually a consequence of the correlation between MAT and growing season length [@kerkhoff_plant_2005; @malhi_productivity_2012; @michaletz_convergence_2014; @michaletz_drivers_2018]. Kerkhoff et al. [-@kerkhoff_plant_2005] and Michaletz et al. [-@michaletz_convergence_2014] find that, within the growing season, there is no significant relationship between primary productivity and MAT, suggesting that the effect of temperature is due to increased length of growing season, rather than an inherent influence of temperature on FACF.

The recent development of the Global Forest Carbon database (ForC), which synthesizes multiple variables and including records of stand history [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018], opens up the possibility for a standardized analysis of global scale variation in multiple FACF and the principle climatic drivers of these patterns. In order to approach this broad topic, we simplify the major gaps in our knowledge to five broad questions and corresponding hypotheses (Table 1). First, we ask how FACF vary with latitude. We then test how these fluxes relate to MAT and MAP, and additionally how they respond to other, less well-studied, climate variables. Finally, we consider the relationship between FACF and seasonality, considering the role of seasonality in explaining variation in carbon fluxes, and the influence of climate on FACF standardized by growing season length. We address the above questions for nine carbon fluxes contained in ForC, allowing for an in-depth exploration of the effect of climate on FACF globally.

@@

We first examined the relationship between latitude and FACF (Q1; Table 1). We tested models with latitude as a linear term (corresponding null: model without latitude) and as a second-order polynomial term (corresponding null: model with latitude as a linear term), and calculated AIC values to determine the best model. Models were accepted as the best model if $\Delta$AIC > 2 with respect to the corresponding null, and were significant with respect to a null model with no fixed term. We also examined relationships among fluxes across latitude, testing whether sums of component fluxes matched the larger fluxes and whether C allocation varied with latitude, as specified below.

To test whether trends in component fluxes across latitude sum to match those of larger fluxes, regression lines for smaller component fluxes were summed to generate new estimates of larger fluxes, which were then compared against the latitudinal regression of the larger flux. Confidence intervals for the larger flux were calculated using the 'bootMer' function from the lme4 package [@bates_fitting_2015]. This analysis was applied to following sets of fluxes: (1) $GPP = NPP + R_{auto}$, (2) $NPP = ANPP + BNPP$, and (3) $ANPP = ANPP_{foliage} + ANPP_{woody.stem}$. In addition, we estimated total belowground C flux (TBCF, not analyzed due to limited data) as $TBCF = BNPP + R_{root}$.

@@

We next examined the relationships of FACF to climate variables (Q2-Q4; Table 1). As with latitude, we tested both linear and polynomial fits for each climate variable. We tested relationships of each FACF (Table 2) against each climate variable (Table S1), but focus presentation--including focal hypotheses (Table 1)--on climate variables that explained >20% of variation in FACF. **Becky, please make this more specific. I drafted this based on the results, but its not very clear.**

To investigate the potential joint and interactive effects of climate variables on carbon fluxes, multivariate models were also specified. We explored a large range of climate variable combinations and ultimately decided to focus on MAT and MAP. We favored these over variables with similar explanatory power because they are the most commonly reported climate variables and are recorded in ForC. MAT was consistently among the best individual predictors of FACF (Table S2), although it was occasionally out-performed by closely correlated climate variables (Fig. S2). In combination with MAT as a primary climate variable, MAP was consistently among the best secondary climate variables. Specifically, compared again a model including MAT x MAP, there were no other variables that, in interaction with MAT, improved AIC by >2.0 and $R^2$ by >0.05. In examinging the joint and interactive effects of climate variables, an additive model including MAT and MAP as fixed effects was accepted when $\Delta$AIC >2 relative to a null including only MAT as a fixed effect. An interactive model including an MAT x MAP interaction was accepted when $\Delta$AIC >2 relative to a null including MAT and MAP as fixed effects.

To test whether and how FACF varied with climate when standardised by growing season length (Q5), we first standardized FACF by dividing by growing season length (as defined above). We then tested for correlations between growing season length-standardised FACF against growing season climate variables. For analyses on data within the growing season, only linear models were specified.

**All analyses were conducted in R (**Version**). Code and results are archived on GitHub....**

@@

All major carbon fluxes decreased linearly with latitude (Fig. 2; Table S2). Latitude was a strong predictor for many of the carbon fluxes--particularly the larger fluxes (Table S2). Specifically, latitude explained 64% of variation in GPP (n = 254, p<0.0001), 50% in NPP (n = 114, p<0.0001) and 45% in ANPP (n = 259, p<0.0001). The FACF that were most poorly predicted by latitude were $BNPP_{root-fine}$ ($R^2$=0.17) and $ANPP_{woody.stem}$ ($R^2$=0.18). For all FACF, the relationship with latitude was best predicted by the linear model.

@@

MAP was a significant (p<0.05) predictor of all FACF but $ANPP_{woody-stem}$ (Figs. 4a, S4-S5; Table S2). However, it explained little variation: with the exception of $R_{auto}$, MAP explained at most 37% of variation in FACF. For the majority of FACF, a polynomial model was the best fit. FACF generally increased with precipitation, up until a saturation point at between 3000 and 4000mm annual precipitation, above which they started to decrease (Figs. 4, S4-S5). The notable exception to this was GPP: the model indicated that GPP continued to increase with precipitation up to measures of at least 5000mm annually (p<0.0001, $R^2$ = 0.33. Data above this point were not available, but the model trend indicated that the saturation point for this model would be around 5000mm MAP.

There was a significant additive effect of MAT and MAP on $GPP$, $ANPP$ and $R_{auto}$ (Fig. 3, Table S3). Accounting for MAT, MAP had a substantial positive effect on $GPP$ and $R_{auto}$ and a small negative effect on $ANPP$. There was a significant interactive effect between MAT and MAP for $NPP$ and $ANPP_{woody-stem}$ (Fig. 3, Table S3). The interaction was negative for $NPP$ and positive for $ANPP_{woody-stem}$. For $ANPP_{foliage}$, $BNPP$, $BNPP_{root-fine}$, and $R_{auto-root}$, MAP did not have a signficant effect when accounting for MAT (Fig. 3, Table S3). For the variables which showed a significant interactive or additive effect between MAT and MAP, no other climate variable, in combination with MAT, significantly improved on that model. {**need to confirm this given changes in MAT MAP results (or you could just drop the sentence.)**}

@@

Temperature seasonality was a significant predictor of FACF. $GPP$, $NPP$, $ANPP$, and $R_{root}$ exhibited a polynomial relationship with seasonality (all p<0.05; Figs. 4e, S6-7; Table S2). $ANPP_{foliage}$, $ANPP_{woody-stem}$ and $R_{auto}$ decreased linearly with temperature seasonality (all p<0.05; Figs. 4e, S6-S7; Table S2). Temperature seasonality was strongly correlated with annual temperature range, which was likewise a similarly strong predictor of FACF (Table S2). FACF were highest where temperature seasonality = 0, and at an annual temperature range of 15$^\circ$C or lower.

@@

Our analysis of a large global database (ForC) reveals how autotrophic carbon fluxes in mature forests vary with latitude and climate on a global scale. We show that, across all nine FACF analyzed, C cycling decreases continually with latitude (*H1.1*; Fig. 2)--a finding that confirms multiple previous studies but contradicts the idea that productivity of temperate forests rivals that of tropical forests [@huston_global_2009]. FACF generally increase in proportion to one another (*H1.2*), with few differences in allocation detectable at this global scale (Fig. S2) and with component fluxes summing appropriately to larger fluxes (Fig. 2), indicating no major, systematic omissions or overestimations of flux components. However, climate explained lower proportions of variability among subsidiary C fluxes (*e.g.*, $ANPP_{woody}$, $BNPP_{fine.root}$, $R_{auto-root}$; Fig. 2; Table S2). Latitudianal variation in FACF is primarily attributable to temperature-related variables (*H3, H4*), particularly MAT (Figs. 3-4). Water availability is also influential, but generally of secondary importance across the climate space occupied by forests (Figs. 3-4). Temperature seasonality and growing season length are closely correlated with MAT and are strong predictors of FACF (*H4*; Figs. 4e-f, S2, S6-S7), though growing season length doesn't improve upon MAT as a predictor. Within the growing season, the influence of climate on C cycling is smaller but still significant for a number of FACF (*H5*; Fig. S9; Table S4). These findings resolve a number of open debates and clarify the big picture of how FACF vary with latitude and climate on a global scale.

Past studies have differed in their conclusions regarding the relationship between FACF and latitude or its correlates (Table 1, *H1*)--quite possibly because of lack of standardization with respect to stand age and disturbance history. Our findings indicate that, among mature, undisturbed stands, FACF are unambiguously highest in the tropical regions, and the relationship is approximately linear (Fig. 2). This contrasts with the suggestion that productivity of temperate forests is similar to that of tropical forests [@huston_global_2009]. Compared to tropical forests, the temperate forest biome has experienced more widespread anthropogenic disturbance and has a larger fraction of secondary stands [@potapov_mapping_2008; @poulter_global_2018], so analyses comparing across latitudinal gradients without controlling for stand age risk confounding age with biome effects. In addition, because carbon allocation varies with stand age [@delucia_forest_2007] (**Doughty et al. 2018: DOI: 10.1111/btp.12504; Anderson-Teixeira et al. 2013: doi: 10.1111/gcb.12194**), age differences may introduce systematic biases into analyses of FACF across latitude or global climatic gradients. For example, woody productivity tends to be higher in rapidly aggrading secondary stands than in old-growth forests, where proportionally more C is allocated to respiration (**Doughty et al. 2018: DOI: 10.1111/btp.12504**);[@delucia_forest_2007; @piao_forest_2010; @kunert_understanding_2019]. By constraining our analysis to stands $\ge$ 100 years old with no record of major recent disturbance, we have clarified the shape of the relationship between FACF and latitude.

>>>>

\newpage
###Abstract
Carbon fixation, allocation, and metabolism by trees set the basis for energy and material flows in forest ecosystems and define their interactions with Earth's changing climate. 
However, we lack a cohesive synthesis on how forest autotrophic carbon fluxes (FACF) vary globally with respect to climate and one another. 
Here, we draw upon # records from the Global Forest Carbon Database (ForC), representing all major forest types and the nine most significant FACF, to comprehensively explore how C cycling in mature, undisturbed forests varies with latitude and climate on a global scale. 
We show that, across all FACF variables analyzed, C cycling decreases linearly with latitude -- a finding that confirms multiple previous studies but contradicts the idea that net primary productivity ($NPP$) of temperate forests rivals that of tropical forests. 
FACF generally displayed similar trends across latitude and multiple climate variables, with few differences in allocation detectable at this global scale, but differed in that latitude explained a lower proportion of variation among subsidiary fluxes (in particular, woody aboveground $NPP$ and belowground $NPP$, $BNPP$).  **BBL: not sure the second part of this sentence merits inclusion in the abstract**
Climate explained a significant proportion (#-#%) of variation in all C fluxes analyzed, with temperature variables in general and mean annual temperature (MAT) in particular being the best predictors of FACF on this global scale. 
The effects of temperature were modified by moisture availability, with FACF reduced under hot and dry conditions and sometimes under very high precipitation. 
FACF declined with temperature seasonality, but growing season length did not improve upon MAT as a predictor.  
Within the growing season, the influence of climate on C cycling was small but significant for a number of FACF. 
These findings clarify how FACF varies with latitude and climate on a global scale. 
As we enter a period of accelerating climatic change, understanding of the fundamental climatic controls on FACF sets a foundation for understanding patterns of change.

@@

Carbon cycling in forests worldwide provides the energetic basis for sustaining the majority of Earth's terrestrial biodiversity and many human populations (**REF**), while strongly influencing atmospheric CO~2~ and climate [@bonan_forests_2008]. Forests' autotrophic carbon fluxes (FACF)--that is, carbon fixation, allocation, and metabolism by trees and other primary producers--sets the energy ultimately available to heterotrophic organisms (including microbes), in turn influencing their abundance (e.g., [Zak et al. 1994](http://esajournals.onlinelibrary.wiley.com/doi/abs/10.2307/1940888); [Niedziałkowska et al. 2010](https://doi.org/10.2980/17-1-3310)) and possibly diversity [@waide_relationship_1999; Chu et al. 2019 (DOI: 10.1111/ele.13175)]. FACF influences all organic matter stocks in forest ecosystems and is linked to cycling of energy, water, and nutrients.

Critically, FACF also define forest interactions with Earth's changing climate. Over 69 Gt of CO~2~ cycle through Earth's forests each year [@badgley_terrestrial_2019], and in recent decades their net C sequestration (~2.4 Gt C yr^-1^) offset roughly 30% of anthropogenic fossil fuel emissions [@pan_large_2011]. As atmospheric carbon dioxide levels continue to rise, driving climate change, forests will play a critical role in shaping the future of Earth's climate [@cavaleri_urgent_2015; @rogelj_mitigation_2018]. However, our ability to draw general macroscopic conclusions regarding global variation in FACF with respect to climate has been limited in that these analyses often mix forests that vary in stand age, disturbance history, and/or management status; do not always sufficiently parse related variables; and typically consider only one or a few variables at a time.

FACF decrease with latitude, but it remains unclear whether and how the shape of this relationship varies between fluxes. FACF are lowest in the boreal regions, and increase into the temperate regions [@luyssaert_co_2007; @huston_global_2009; @beer_terrestrial_2010; @piao_forest_2010; @jung_global_2011]. However, evidence is inconclusive on whether primary productivity continues to increase into the tropics. Evidence for this is further complicated by the fact that different studies use different measures of productivity to explore these relationships. For example, modelling of global terrestrial ecosystem gross primary productivity ($GPP$) through upscaling and calibration of eddy flux measurements indicates that GPP peaks in tropical forests [@beer_terrestrial_2010; @jung_global_2011; @badgley_terrestrial_2019; @li_mapping_2019]. In contrast, some studies suggest that the highest values of net primary productivity ($NPP$) may be found in temperate forests [@luyssaert_co_2007; @huston_global_2009], while others find $NPP$ highest in the tropics and decreasing with latitude [@simova_enigma_2017]. Other studies have chosen to focus exclusively on above-ground net primary productivity ($ANPP$), finding evidence of a weak negative relationship between $ANPP$ and latitude [@huston_global_2009; @gillman_latitude_2015].

The latitudinal gradient in FACF is primarily driven by climate, which is a significant driver of FACF across broad spatial scales [@luyssaert_co_2007; @cleveland_relationships_2011; @hursh_sensitivity_2017]. The majority of studies have focused on exploring the relationships of FACF to mean annual temperature (MAT) and mean annual precipitation (MAP), as the most commonly reported site-level climate variables. There is strong evidence that both MAT and MAP show significant positive relationships with FACF [@chu_does_2016]. However, as with latitude, the shape of those relationships is not always clear, and, again, is complicated by the use of different measures of FACF across studies. Various measures of primary productivity {**FACF?**} saturate at high levels of MAP, though the saturation points identified vary from 1500mm [@luyssaert_co_2007] up to 2445mm MAP [@schuur_productivity_2003]. Studies of the influence of MAT on productivity {**FACF?**} are less conclusive. Luyssaert et al. [-@luyssaert_co_2007] examined GPP and NPP and found that, while $GPP$ increases linearly with MAT, NPP saturates at around 10$^\circ$C MAT. In contrast, Larjavaara and Muller-Landau [-@larjavaara_temperature_2012], find that increases in $GPP$ saturate at approximately 25$^\circ$C MAT, while Schuur [-@schuur_productivity_2003] shows that $NPP$ increases linearly with temperature. Taylor et al. [-@taylor_temperature_2017] showed a positive interaction between MAT and MAP in shaping tropical forest productivity, such that high rainfall had a negative effect on productivity in cooler climates, compared to a positive effect in warmer climates. Such complicated dynamics play out with belowground fluxes as well, with typically dominant soil temperature mediated by moisture and carbon supply (Hursh et al. 2017, http://dx.doi.org/10.1111/gcb.13489; Xu et al. 2016 http://dx.doi.org/10.1016/j.jplph.2016.08.007).

MAT and MAP are very coarse measures of climate, and so fail to capture much variation in climate on an intra-annual scale, including the effects of factors such as growing season length, number of frost-free days, temperature seasonality, and dry season length. Some studies have suggested that the apparently strong relationship between MAT and FACFs is actually a consequence of the correlation between MAT and growing season length [@kerkhoff_plant_2005; @malhi_productivity_2012; @michaletz_convergence_2014; @michaletz_drivers_2018]. Kerkhoff et al. [-@kerkhoff_plant_2005] and Michaletz et al. [-@michaletz_convergence_2014] find that, within the growing season, there is no significant relationship between primary productivity and MAT, suggesting that the effect of temperature is due to increased length of growing season, rather than an inherent influence of temperature on FACF.

In addition, FACF can be influenced by other, often interactive, factors across a range of scales [@cleveland_relationships_2011]. On a local scale, stand age [@litton_carbon_2007; @gillman_latitude_2015], biodiversity [@liang_positive_2016], management [@simova_enigma_2017], nutrient availability [@aragao_above_2009], and altitude [@girardin_net_2010; @malhi_variation_2017] all impact FACF. There is evidence that FACFs also respond to variables such as cloud cover [@taylor_temperature_2017], solar radiation [@fyllas_solar_2017], and potential evapotranspiration [@kerkhoff_plant_2005].

The recent development of the Global Forest Carbon database (ForC), which synthesizes multiple variables and including records of stand history [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018], opens up the possibility for a standardized analysis of global scale variation in multiple FACF and the principle climatic drivers of these patterns. In order to approach this broad topic, we simplify the major gaps in our knowledge to five broad questions and corresponding hypotheses (Table 1). First, we ask how nine FACF in ForC vary with latitude. We then test how these fluxes relate to MAT and MAP, and additionally how they respond to other, less well-studied, climate variables. Finally, we consider the relationship between FACF and seasonality, considering the role of seasonality in explaining variation in carbon fluxes, and the influence of climate on FACF standardized by growing season length.

@@

We first examined the relationship between latitude and FACF (Q1; Table 1). We tested models with latitude as a linear term (corresponding null: model without latitude) and as a second-order polynomial term (corresponding null: model with latitude as a linear term), and calculated AIC values to determine the best model. **BBL: careful about terminology here. In statistics a "linear model" includes polynomial models, which are linear in their parameters. Thus I (and many readers perhaps) find this confusing.** Models were accepted as the best model if $\Delta$AIC > 2 with respect to the corresponding null, and were significant with respect to a null model with no fixed term.

To test whether trends in component fluxes across latitude sum to match those of larger fluxes, regression lines for smaller component fluxes were summed to generate new estimates of larger fluxes, which were then compared against the latitudinal regression of the larger flux. Confidence intervals for the larger flux were calculated using the 'bootMer' function from the lme4 package [@bates_fitting_2015]. This analysis was applied to the following sets of fluxes: (1) $GPP = NPP + R_{auto}$, (2) $NPP = ANPP + BNPP$, and (3) $ANPP = ANPP_{foliage} + ANPP_{woody.stem}$. In addition, we estimated total belowground C flux (TBCF, not analyzed due to limited data) as $TBCF = BNPP + R_{root}$.

@@

We next examined the relationships of FACF to climate variables (Q2-Q4; Table 1). As with latitude, we tested both linear and polynomial fits for each climate variable. We tested relationships of each FACF (Table 2) against each climate variable (Table S1), but focus presentation--including focal hypotheses (Table 1)--on climate variables that explained >20% of variation in FACF. **Becky, please make this more specific. I drafted this based on the results, but its not very clear.** **BBL: I agree, not clear and might trigger reviewer concerns about tailoring hypotheses to significance.**

Multivariate models were used to investigate the potential joint and interactive effects of climate variables on carbon fluxes. We explored a large range of climate variable combinations and ultimately decided to focus on MAT and MAP. We favored these over variables with similar explanatory power because they are the most commonly reported climate variables. MAT was consistently among the best individual predictors of FACF (Table S2), although it was occasionally out-performed by closely correlated climate variables (Fig. S2). In combination with MAT as a primary climate variable, MAP was consistently among the best secondary climate variables. Specifically, compared against a model including MAT x MAP, there were no other variables that, in interaction with MAT, improved AIC by >2.0 and $R^2$ by >0.05. In examining the joint and interactive effects of climate variables, an additive model including MAT and MAP as fixed effects was accepted when $\Delta$AIC >2 relative to a null including only MAT as a fixed effect. An interactive model including an MAT x MAP interaction was accepted when $\Delta$AIC >2 relative to a null including MAT and MAP as fixed effects.

To test whether and how FACF varied with climate when standardised by growing season length (Q5), we first standardized FACF by dividing by growing season length (as defined above). We then tested for correlations between these standardised fluxes and growing season climate variables, using only linear models.

**All analyses were conducted in R (**Version**). Code and data necessary to reproduce all results are archived on GitHub....**

@@

All major carbon fluxes decreased linearly with latitude (Fig. 2; Table S2). Latitude was a strong predictor for many of the carbon fluxes--particularly the larger fluxes (Table S2). Specifically, latitude explained 64% of variation in GPP (n = 254, p<0.0001), 50% in NPP (n = 114, p<0.0001) and 45% in ANPP (n = 259, p<0.0001). The FACF that were most poorly predicted by latitude were $BNPP_{root-fine}$ ($R^2$=0.17) and $ANPP_{woody.stem}$ ($R^2$=0.18). For all FACF, the relationship with latitude was best predicted by the first-order linear model.

@@

MAP was a significant (p<0.05) predictor of all FACF except $ANPP_{woody-stem}$ (Figs. 4a, S4-S5; Table S2). However, it explained little variation: with the exception of $R_{auto}$, MAP explained at most 37% of variation in FACF. For the majority of FACF, a polynomial model **BBL: of what order?** provided the best fit. FACF generally increased with precipitation, up until a saturation point at between 3000 and 4000mm annual precipitation, above which they started to decrease (Figs. 4, S4-S5). The notable exception to this was GPP: the model indicated that GPP continued to increase with precipitation up to measures of at least 5000mm annually (p<0.0001, $R^2$ = 0.33. Data above this point were not available, but the model trend suggested a saturation point around 5000mm MAP.

There was a significant additive effect of MAT and MAP on $GPP$, $ANPP$ and $R_{auto}$ (Fig. 3, Table S3). Accounting for MAT, MAP had a substantial positive effect on $GPP$ and $R_{auto}$ and a small negative effect on $ANPP$. There was a significant interactive effect between MAT and MAP for $NPP$ and $ANPP_{woody-stem}$ (Fig. 3, Table S3). The interaction was negative for $NPP$ and positive for $ANPP_{woody-stem}$. For $ANPP_{foliage}$, $BNPP$, $BNPP_{root-fine}$, and $R_{auto-root}$, MAP did not have a signficant effect when accounting for MAT (Fig. 3, Table S3). For the variables which showed a significant interactive or additive effect between MAT and MAP, no other climate variable, in combination with MAT, significantly improved on that model. {**need to confirm this given changes in MAT MAP results (or you could just drop the sentence.)**}
**BBL: this paragraph is pretty dense. Perhaps distill to a summary sentence and move rest to SI?**

@@

Temperature seasonality was a significant predictor of FACF. $GPP$, $NPP$, $ANPP$, and $R_{root}$ exhibited a polynomial relationship with seasonality (all p<0.05; Figs. 4e, S6-7; Table S2). $ANPP_{foliage}$, $ANPP_{woody-stem}$ and $R_{auto}$ decreased linearly with temperature seasonality (all p<0.05; Figs. 4e, S6-S7; Table S2). Temperature seasonality was strongly correlated with annual temperature range, which was likewise a similarly strong predictor of FACF (Table S2). FACF were highest where temperature seasonality = 0, and at an annual temperature range of 15$^\circ$C or lower. **BBL: perhaps put this into an ecosystem context; what are these? Aseasonal subtropical places?**

@@

Our analysis of a large global database (ForC) reveals how autotrophic carbon fluxes in mature forests vary with latitude and climate on a global scale. We show that, across all nine FACF analyzed, C cycling decreases continually with latitude (*H1.1*; Fig. 2)--a finding that confirms multiple previous studies but contradicts the idea that productivity of temperate forests rivals that of tropical forests [@huston_global_2009]. FACF generally increase in proportion to one another (*H1.2*), with few differences in allocation detectable at this global scale (Fig. S2) and with component fluxes summing appropriately to larger fluxes (Fig. 2), indicating no major, systematic omissions or overestimations of flux components. However, climate explained lower proportions of variability among subsidiary C fluxes (*e.g.*, $ANPP_{woody}$, $BNPP_{fine.root}$, $R_{auto-root}$; Fig. 2; Table S2).

Latitudinal variation in FACF is **BBL: be consistent in tense** primarily attributable to temperature-related variables (*H3, H4*), particularly MAT (Figs. 3-4). Water availability is also influential, but generally of secondary importance across the climate space occupied by forests (Figs. 3-4). Temperature seasonality and growing season length are closely correlated with MAT and are strong predictors of FACF (*H4*; Figs. 4e-f, S2, S6-S7), though growing season length did not improve upon MAT as a predictor. Within the growing season, the influence of climate on C cycling is smaller but still significant for a number of FACF (*H5*; Fig. S9; Table S4).

These findings clarify how FACF vary with latitude and climate on a global scale. Past studies have differed in their conclusions regarding the relationship between FACF and latitude or its correlates (Table 1, *H1*)--quite possibly because of lack of standardization with respect to stand age and disturbance history. Our findings indicate that, among mature, undisturbed stands, FACF are unambiguously highest in the tropical regions, and the relationship is approximately linear (Fig. 2). This contrasts with the suggestion that productivity of temperate forests is similar to that of tropical forests [@huston_global_2009]. Compared to tropical forests, the temperate forest biome has experienced more widespread anthropogenic disturbance and has a larger fraction of secondary stands [@potapov_mapping_2008; @poulter_global_2018], so analyses comparing across latitudinal gradients without controlling for stand age risk confounding age with biome effects. In addition, because carbon allocation varies with stand age [@delucia_forest_2007] (**Doughty et al. 2018: DOI: 10.1111/btp.12504; Anderson-Teixeira et al. 2013: doi: 10.1111/gcb.12194**), age differences may introduce systematic biases into analyses of FACF across latitude or global climatic gradients. For example, woody productivity tends to be higher in rapidly aggrading secondary stands than in old-growth forests, where proportionally more C is allocated to respiration (**Doughty et al. 2018: DOI: 10.1111/btp.12504**);[@delucia_forest_2007; @piao_forest_2010; @kunert_understanding_2019].

We show that FACF are broadly consistent in their responses to climate drivers on the global scale, with at most modest trends in C allocation among the variable pairs tested (Figs. 2, S3). 
This parallels the observation that C allocation across multiple FACF varies little with respect to climate along a steep tropical elevational gradient [@malhi_variation_2017; but see @moser_elevation_2011], 
and is not surprising given that trees face universal constraints in carbon allocation (**REFS**). 
The one trend in allocation that emerged from our analysis was a tendency for greater C allocation aboveground--and to foliage in particular--in warm tropical climates (Fig. S3). This is consistent with observations of increasing allocation to roots with declining temperature across a tropical elevational gradient [@moser_elevation_201], 
and with observations and theory predicting relatively higher belowground carbon allocation at higher latitudes [@gill_belowground_2016].
It is also consistent with findings that as forest productivity increases, proportionaly more carbon is allocated to $ANPP_{foliage}$ relative to $BNPP_{root-fine}$ (*Chen et al. 2019; doi: 10.1111/ele.13193*) or $ANPP_{woody}$ [@hofhansl_new_2015]. **( check / comment on  @litton_carbon_2007 )**

One previously hypothesized trend that was not supported by our analysis was that tropical forests tend to have low carbon use efficiency ($CUE$= $NPP$/$GPP$=($GPP$-$R_{auto}$/$GPP$), which is based on observations of low $CUE$ in old-growth tropical forests relative to (mostly younger) extratropical forests [@delucia_forest_2007; @malhi_productivity_2012; @anderson-teixeira_carbon_2016]. Our analysis, limited to mature forests, shows no such trend (Fig. S3). $CUE$ is known to decline with forest age [@delucia_forest_2007; @piao_forest_2010; @collalti_is_2019], but appears to be roughly independent of $GPP$ [@litton_carbon_2007]. (*This finding may have some important implications for modeling and our broader understanding. I'm not sure offhand how much theory has been built on the idea of low CUE in tropical forests because of warm temperatures...*)
